Allocation Removal by Partial Evaluation in a Tracing JIT

### 2010-10-13 talk/pepm2011/paper.tex
re-import benchmark times, fix one XXX

\begin{figure*}
\begin{tabular}{|l||r|rr|rr|rr|rr|}
\hline
	&num loops	&new	&removed	&get/set	&removed &guard &removed	&all ops	&removed\\
\hline
crypto\_pyaes	&78	&3088	&50\%	&57148	&25\%	&9055	&95\%	&137189	&80\%\\
django	&51	&673	&54\%	&19318	&18\%	&3876	&93\%	&55682	&85\%\\
fannkuch	&43	&171	&49\%	&886	&63\%	&1159	&81\%	&4935	&45\%\\
go	&517	&12234	&76\%	&200842	&21\%	&53138	&90\%	&568542	&84\%\\
html5lib	&498	&14432	&68\%	&503390	&11\%	&71592	&94\%	&1405780	&91\%\\
meteor-contest	&59	&277	&36\%	&4402	&31\%	&1078	&83\%	&12862	&68\%\\
nbody	&13	&96	&38\%	&443	&69\%	&449	&78\%	&2107	&38\%\\
pyflate-fast	&162	&2278	&55\%	&39126	&20\%	&8194	&92\%	&112857	&80\%\\
raytrace-simple	&120	&3118	&59\%	&91982	&15\%	&13572	&95\%	&247436	&89\%\\
richards	&87	&844	&4\%	&49875	&22\%	&4130	&91\%	&133898	&83\%\\
spambayes	&314	&5608	&79\%	&117002	&11\%	&25313	&94\%	&324125	&90\%\\
spectral-norm	&38	&360	&64\%	&5553	&20\%	&1122	&92\%	&11878	&77\%\\
telco	&46	&1257	&90\%	&37470	&3\%	&6644	&99\%	&98590	&97\%\\
twisted-names	&214	&5273	&84\%	&100010	&10\%	&23247	&96\%	&279667	&92\%\\
\hline
total	&2240	&49709	&70\%	&1227447	&14\%	&222569	&93\%	&3395548	&89\%\\
\hline
\end{tabular}
\caption{Number of Operations and Percentage Removed By Optimization}
\label{fig:numops}
\end{figure*}

In addition to the count of operations we also performed time measurements.
All benchmarks were run 50 times in the same process, to give the JIT time to produce machine
code. The arithmetic mean of the times of the last 30 runs were used as the
result. The errors were computed using a confidence interval with a 95\%
confidence level \cite{georges_statistically_2007}. The results are reported in Figure~\ref{fig:times}.
With the optimization turned on, PyPy's Python interpreter outperforms CPython
in all benchmarks except spambayes (which heavily relies on regular expression
performance XXX). All benchmarks are improved by the allocation removal
optimization, some as much as XXX. XXX Psyco

@@

\begin{figure*}
\begin{tabular}{lrrrr}
		&CPython	&JIT no optimizations	&JIT allocation removal	&JIT full\\
crypto\_pyaes	&1958.40 $\pm$ 0.32	&782.14 $\pm$ 3.08	&150.86 $\pm$ 3.62	&125.68 $\pm$ 3.02\\
django		&630.91 $\pm$ 0.13	&365.90 $\pm$ 1.57	&124.41 $\pm$ 0.88	&117.36 $\pm$ 0.91\\
fannkuch	&1344.15 $\pm$ 0.36	&378.18 $\pm$ 0.36	&309.04 $\pm$ 1.56	&299.12 $\pm$ 0.25\\
go		&595.31 $\pm$ 0.51	&1067.26 $\pm$ 12.20	&121.76 $\pm$ 4.15	&125.57 $\pm$ 3.90\\
html5lib	&8678.67 $\pm$ 23.37	&17665.16 $\pm$ 2695.45	&6739.95 $\pm$ 1225.25	&6699.24 $\pm$ 1295.92\\
meteor-contest	&241.53 $\pm$ 0.09	&291.11 $\pm$ 0.86	&274.09 $\pm$ 0.37	&272.77 $\pm$ 0.30\\
nbody		&396.55 $\pm$ 0.24	&84.77 $\pm$ 0.14	&69.01 $\pm$ 0.08	&68.71 $\pm$ 0.09\\
pyflate-fast	&1991.33 $\pm$ 1.90	&1818.70 $\pm$ 4.30	&1062.03 $\pm$ 4.56	&1052.02 $\pm$ 3.24\\
raytrace-simple	&1598.84 $\pm$ 1.07	&1126.93 $\pm$ 11.42	&93.45 $\pm$ 4.75	&92.17 $\pm$ 4.15\\
richards	&216.47 $\pm$ 0.20	&169.84 $\pm$ 1.21	&11.81 $\pm$ 0.09	&11.90 $\pm$ 0.29\\
spambayes	&193.30 $\pm$ 0.03	&331.37 $\pm$ 14.93	&238.83 $\pm$ 6.10	&237.23 $\pm$ 5.97\\
spectral-norm	&375.91 $\pm$ 0.08	&221.24 $\pm$ 0.07	&31.96 $\pm$ 0.06	&28.79 $\pm$ 0.09\\
telco		&789.67 $\pm$ 1.14	&622.67 $\pm$ 1.86	&155.33 $\pm$ 1.82	&154.67 $\pm$ 1.82\\
twisted\_names	&6.61 $\pm$ 0.00	&8.85 $\pm$ 0.09	&4.35 $\pm$ 0.05 &4.35 $\pm$ 0.04\\
\end{tabular}
\caption{Benchmark Times}
\label{fig:times}
\end{figure*}

>>>>

\begin{figure*}
\begin{center}
\begin{tabular}{|l||r|rr|rr|rr|rr|}
\hline
	&num loops	&new	&removed	&get/set	&removed &guard &removed	&all ops	&removed\\
\hline
crypto\_pyaes	&78	&3088	&50\%	&57148	&25\%	&9055	&95\%	&137189	&80\%\\
django	&51	&673	&54\%	&19318	&18\%	&3876	&93\%	&55682	&85\%\\
fannkuch	&43	&171	&49\%	&886	&63\%	&1159	&81\%	&4935	&45\%\\
go	&517	&12234	&76\%	&200842	&21\%	&53138	&90\%	&568542	&84\%\\
html5lib	&498	&14432	&68\%	&503390	&11\%	&71592	&94\%	&1405780	&91\%\\
meteor-contest	&59	&277	&36\%	&4402	&31\%	&1078	&83\%	&12862	&68\%\\
nbody	&13	&96	&38\%	&443	&69\%	&449	&78\%	&2107	&38\%\\
pyflate-fast	&162	&2278	&55\%	&39126	&20\%	&8194	&92\%	&112857	&80\%\\
raytrace-simple	&120	&3118	&59\%	&91982	&15\%	&13572	&95\%	&247436	&89\%\\
richards	&87	&844	&4\%	&49875	&22\%	&4130	&91\%	&133898	&83\%\\
spambayes	&314	&5608	&79\%	&117002	&11\%	&25313	&94\%	&324125	&90\%\\
spectral-norm	&38	&360	&64\%	&5553	&20\%	&1122	&92\%	&11878	&77\%\\
telco	&46	&1257	&90\%	&37470	&3\%	&6644	&99\%	&98590	&97\%\\
twisted-names	&214	&5273	&84\%	&100010	&10\%	&23247	&96\%	&279667	&92\%\\
\hline
total	&2240	&49709	&70\%	&1227447	&14\%	&222569	&93\%	&3395548	&89\%\\
\hline
\end{tabular}
\end{center}
\caption{Number of Operations and Percentage Removed By Optimization}
\label{fig:numops}
\end{figure*}

In addition to the count of operations we also performed time measurements.
All benchmarks were run 50 times in the same process, to give the JIT time to produce machine
code. The arithmetic mean of the times of the last 30 runs were used as the
result. The errors were computed using a confidence interval with a 95\%
confidence level \cite{georges_statistically_2007}. The results are reported in Figure~\ref{fig:times}.
With the optimization turned on, PyPy's Python interpreter outperforms CPython
in all benchmarks except spambayes (which heavily relies on regular expression
performance and thus is not helped much by our Python JIT) and meteor-contest.
All benchmarks are improved by the allocation removal optimization, some as much
as XXX. XXX Psyco

@@

\begin{figure*}
\begin{center}
\begin{tabular}{|l||r|r|r|r|}
\hline
	&Cpython	&Psyco	&PyPy w/o optimizations	&PyPy w/ optimizations \\
\hline
crypto\_pyaes	&2757.80 $\pm$ 0.98	&67.90 $\pm$ 0.47	&1652.00 $\pm$ 4.00 &266.86 $\pm$ 5.94 \\
django	&993.19 $\pm$ 0.50	&913.51 $\pm$ 4.22	&694.73 $\pm$ 2.86 &259.53 $\pm$ 1.79 \\
fannkuch	&1987.22 $\pm$ 2.02	&944.44 $\pm$ 0.61	&566.99 $\pm$ 1.06 &466.87 $\pm$ 1.85 \\
go	&947.21 $\pm$ 1.58	&445.96 $\pm$ 0.68	&2197.71 $\pm$ 25.21 &316.15 $\pm$ 9.33 \\
html5lib	&13987.12 $\pm$ 19.51	&17398.25 $\pm$ 36.50	&27194.45 $\pm$ 46.62	&10092.19 $\pm$ 23.50 \\
meteor-contest	&346.98 $\pm$ 0.35	&215.66 $\pm$ 0.23	&433.04 $\pm$ 1.45 &392.85 $\pm$ 0.87 \\
nbody\_modified	&637.90 $\pm$ 1.82	&256.78 $\pm$ 0.18	&135.55 $\pm$ 0.33 &103.93 $\pm$ 0.25 \\
pyflate-fast	&3169.35 $\pm$ 1.89	&1278.16 $\pm$ 3.13	&3285.89 $\pm$ 8.51	&1822.36 $\pm$ 11.52 \\
raytrace-simple	&2744.60 $\pm$ 51.72	&1072.66 $\pm$ 1.08	&2778.27 $\pm$ 15.13	&647.24 $\pm$ 5.44 \\
richards	&354.06 $\pm$ 1.00	&63.48 $\pm$ 0.15	&383.93 $\pm$ 3.28 &88.32 $\pm$ 0.91 \\
spambayes	&299.16 $\pm$ 0.35	&338.68 $\pm$ 3.14	&580.90 $\pm$ 24.68 &397.37 $\pm$ 10.60 \\
spectral-norm	&478.63 $\pm$ 0.80	&139.83 $\pm$ 1.54	&353.51 $\pm$ 1.39 &112.10 $\pm$ 1.17 \\
telco	&1207.67 $\pm$ 2.03	&730.00 $\pm$ 2.66	&1296.08 $\pm$ 4.37 &495.23 $\pm$ 2.14 \\
twisted\_names	&9.58 $\pm$ 0.01	&10.43 $\pm$ 0.01	&17.99 $\pm$ 0.27	&7.13 $\pm$ 0.09 \\
\hline
\end{tabular}
\end{center}
\caption{Benchmark Times}
\label{fig:times}
\end{figure*}
